Assignment 1 Reproducible research
==================================

##Loading and processing the data

###Let's first load the data and see how it looks like
```{r,echo=FALSE}
###Load the csv data
raw_data <- read.csv(unzip("repdata-data-activity.zip"))
head(raw_data)
```
###Now, let's transform the date column using the function as.Date()
```{r,warning=FALSE,message=FALSE}
###Transform the data: create 3 new variables: year, month and day
library(dplyr)
data<-raw_data
data<-mutate(data,date=as.Date(date))
head(data)
```

##What is mean total number of steps taken per day?

###Calculating the total number of steps per day
```{r,warning=FALSE,message=FALSE}
steps_per_day<-data %>%
        group_by(date)%>%
        summarise(sum=sum(steps,na.rm=TRUE))
head(steps_per_day)
```
###Histogram of the total number of steps taken each day
```{r,warning=FALSE,message=FALSE}
library(ggplot2)
h<-ggplot(steps_per_day, aes(date,sum))
h+geom_histogram(stat="identity")
```
###Calculating and reporting the mean and median of the total number of steps taken per day
```{r}
mean(steps_per_day$sum)
median(steps_per_day$sum)
```

##What is the average daily activity pattern?

###Time series plot of the 5-minute interval and the average number of steps taken, averaged across all days
```{r}
data<-data %>% mutate(interval_in_day=interval %% 1440)
data_time_series <- data %>%
        group_by(interval_in_day) %>%
        summarise(mean_steps=mean(steps,na.rm=TRUE))
plot(data_time_series$interval_in_day, data_time_series$mean_steps,type="l")
head(data_time_series)
```
###The 5-minute Interval that contains the maximum number of steps in average is
```{r}
data_time_series$interval_in_day[which(data_time_series$mean_steps==max(data_time_series$mean_steps))]
```

##Imputing missing values

###Let's calculate and report the total number of missing values in the dataset
```{r}
number_rows_na <- nrow(data)-sum(complete.cases(data))
number_rows_na
```
###Strategy for filling in all of the missing values in the dataset
Chosen strategy:We will replace the missing value with the mean of the 5-minute interval
We will also create a new dataset with the missing values replated by the corresponding value
```{r}
data2<-data
for (i in 1:nrow(data2)){
        if (is.na(data2[i,1])){
                data2[i,1]<-data_time_series$mean_steps[which(data2[i,4]==data_time_series$interval)]
        }
}
head(data2)
```
###Make a histogram of the total number of steps taken each day 
```{r}
summary_data2 <- data2 %>%
        group_by(date) %>%
        summarise(sum=sum(steps))
h2 <- ggplot(summary_data2,aes(date,sum))
h2+ geom_histogram(stat="identity")
```
###Calculating and reporting the mean and median total number of steps taken per day
```{r}
mean(summary_data2$sum)
median(summary_data2$sum)
```
We notice that the mean and the median are different from those calculated in the first part of the assignment and that the mean and the median are equal now.

##Are there differences in activity patterns between weekdays and weekends?

###Creating a new factor variable in the dataset with two levels - "weekday" and "weekend"
```{r}
data2<-mutate(data2,weekday=weekdays(date))
data2[data2$weekday=="saturday"|data2$weekday=="sunday",5]<-"weekend"
data2[!(data2$weekday=="saturday"|data2$weekday=="sunday"),5]<-"weekday"
head(data2)

```
###Make a panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days
```{r}
summary_weekday_weekend <- data2 %>%
        group_by(weekday,interval_in_day) %>%
        summarize(mean=round(mean(steps),0))
head(summary_weekday_weekend)
par(mfrow=c(2,1))
weekday<-filter(summary_weekday_weekend,weekday=="weekday")
weekend<-filter(summary_weekday_weekend,weekday=="weekend")
plot(weekday$interval_in_day,weekday$mean,type="l")
plot(weekend$interval_in_day,weekend$mean,type="l")
